---
title: "Startup Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load data

The startup data consists of multiple files

Bellow is the loading of the data.

```{r message=FALSE}

library(readr)

#People
people = read_csv("./people.csv")
(peep_n = colnames(people))
peep_n[2] = "person_id"
colnames(people) = peep_n
attr(people, "spec") = NULL
str(people)

```

```{r message=FALSE}
#Company
companies = read_csv("./objects.csv")
#(comp_n = colnames(companies))
attr(companies, "spec") = NULL
attr(companies, "problems") = NULL
colnames(companies)[1] = "company_id"
str(companies)
```

```{r message=FALSE}

degrees = read_csv("./degrees.csv")
colnames(degrees)
colnames(degrees)[2] = "person_id"
colnames(degrees)
attr(degrees, "spec") = NULL
str(degrees)

```


```{r message=FALSE}
comp_peep = read_csv("./relationships.csv")
colnames(comp_peep)[3] = "person_id"
colnames(comp_peep)[4] = "company_id"
colnames(comp_peep)
attr(comp_peep, "spec") = NULL
str(comp_peep)
```


```{r message=FALSE}
#install.packages("reshape")
library(reshape)
library(data.table)

invest_rounds = read_csv("./funding_rounds.csv")
attr(invest_rounds, "spec") = NULL
colnames(invest_rounds)[3]="company_id"
str(invest_rounds)

```


```{r message=FALSE}
aquisitions = read_csv("./acquisitions.csv")
colnames(aquisitions)[4] = "company_id"
colnames(aquisitions)[6] = "aquisition_price_amount"

attr(aquisitions, "spec") = NULL
str(aquisitions)

#unique(aquisitions$price_currency_code)
#aggregate(aquisitions[aquisitions$price_currency_code == "USD",c("price_currency_code", "company_id")], by = list(aquisitions$price_currency_code), FUN = counts)#aquisitions = 

```

# Data Preparation

## Company Founders


This joins the people, degrees and their relationship to the companies. Also,this section "normalizes" the data, particularly the discreet data, so it can be used for regression.

```{r}
full_peep = merge(people, degrees, by="person_id")
str(full_peep)
```

```{r }
#join people with comp_peeps
founders = merge(full_peep, comp_peep, by="person_id")
str(founders)
#View(founders)
```

```{r}
#Identify founders

#titles = unique(founders$title)
#View(titles)
#subjects = unique(founders$subject)
#institutions = unique(founders$institution)
#c = unique(founders$company_id)
#deg = unique(founders$degree_type)

#length(titles)
#length(subjects)
#length(institutions)
#length(c)
#length(deg)

#length(titles[grep("founder", ignore.case = TRUE, x = titles)])
#length(titles[grep("ceo", ignore.case = TRUE, x = titles)])

#get rows with founder, ceos, chairmans only
leads = c(grep("founder", ignore.case = TRUE, x = founders$title), grep("ceo", ignore.case = TRUE, x = founders$title), grep("chairman", ignore.case = TRUE, x = founders$title))
founders[grep("founder", ignore.case = TRUE, x = founders$title), "title"] = "Founder"
founder_leads = founders[leads, ]
#nrow(founder_leads)



#get rows with board
#advisors = c(grep("board", ignore.case = TRUE, x = founders$title), grep("advisor", ignore.case = TRUE, x = founders$title))
#founder_advisors = founders[advisors, ]
#nrow(founder_advisors)

#paste("Companies with founders: ", length(unique(founder_leads$company_id)))

#founder_lead_advisor = founder_leads[founder_leads$title %in% founder_advisors$title, ]
#paste("Companies with founders and advisors: ", length(unique(founder_lead_advisor$company_id)))



#Functions to coubt rows.=
count_disticts = function(x){
  length(unique(x))
}

counts = function(x){
  length(x)
}
```


"Normalize" the title information for the founders' degree type so that it can be used as a potential regression predictor.

```{r}
founder_leads_clean = founder_leads
founder_leads_clean[grep("MS ", ignore.case = TRUE, x = founder_leads_clean$degree_type), c("degree_type")] = "MS"
founder_leads_clean[grep("Master in Business|Master of Business|M\\.B\\.A\\.|MBA|Executive", ignore.case = TRUE, x = founder_leads_clean$degree_type), c("degree_type")] = "MBA"
founder_leads_clean[grep("Ph|doc|pos.*grad", ignore.case = TRUE, x = founder_leads_clean$degree_type), c("degree_type")] = "Ph.D."
founder_leads_clean[grep("Bachelor of Science|BSc|B\\.S|B\\.E|BE|BS|Ing|Eng|Bsc|SCB", ignore.case = TRUE, x = founder_leads_clean$degree_type), c("degree_type")] = "BS"
founder_leads_clean[grep("Bachelor|BA|B\\.A", ignore.case = TRUE, x = founder_leads_clean$degree_type), c("degree_type")] = "BA"
founder_leads_clean[grep("B[A-Z][A-Z]+|B\\.|BC", ignore.case = FALSE, x = founder_leads_clean$degree_type), c("degree_type")] = "Bachelors"
founder_leads_clean[grep("MS|M\\.S", ignore.case = TRUE, x = founder_leads_clean$degree_type), c("degree_type")] = "MS"
founder_leads_clean[grep("Master|Graduate|Fellow", ignore.case = TRUE, x = founder_leads_clean$degree_type), c("degree_type")] = "Masters"
founder_leads_clean[grep("M[A-Z]+|M.", ignore.case = FALSE, x = founder_leads_clean$degree_type), c("degree_type")] = "Masters"
founder_leads_clean[grep("No.*degre|high.*school|secondary educa|no.*complete|course|incomplete|none|did not|prepa|[0-9].*[0-9]+ Grade", ignore.case = TRUE, x = founder_leads_clean$degree_type), c("degree_type")] = "No Degree"
founder_leads_clean[grep("Law|JD|J\\.D\\.", ignore.case = FALSE, x = founder_leads_clean$degree_type), c("degree_type")] = "Law"
founder_leads_clean[grep("Business", ignore.case = TRUE, x = founder_leads_clean$degree_type), c("degree_type")] = "MBA"
founder_leads_clean[grep("Undergr|Diplom|degree|college|under grad|lic|dipl", ignore.case = TRUE, x = founder_leads_clean$degree_type), c("degree_type")] = "Bachelors"

founder_leads_clean[!(founder_leads_clean$degree_type %in% c("MS", "MBA", "Ph.D.", "BS", "BA", "Bachelors", "MS", "Masters", "No Degree", "Law", "MBA")), c("degree_type")] = "Degree"

#View(unique(founder_leads_clean$degree_type))
analysis = aggregate(founder_leads_clean[,c("degree_type", "person_id", "company_id")], by = list(founder_leads_clean$degree_type), FUN = counts)


#analysis[order(-analysis$person_id), ]
#analysis[order(analysis$person_id)]
founder_leads_clean = na.omit(founder_leads_clean[,c("degree_type", "subject", "institution", "company_id", "title")])
founder_leads_clean = founder_leads_clean[order(founder_leads_clean$company_id),]
#View(founder_leads_clean)

```

Reshape the people subdataset into a wider table that has founder information per company, each in a row of data.

```{r}

founders_bg = cbind(leader1_degree_type = rep(0, nrow(founder_leads_clean)), leader1_subject = rep(0, nrow(founder_leads_clean)), leader1_institution = rep(0, nrow(founder_leads_clean)), leader2_degree_type = rep(0, nrow(founder_leads_clean)), leader2_subject = rep(0, nrow(founder_leads_clean)), leader2_institution = rep(0, nrow(founder_leads_clean)), founders_ceos_chairmans = rep(0, nrow(founder_leads_clean)))

nleads = 2
prev_company = "c"
company_index = 0
l = 0
for(i in  1:nrow(founder_leads_clean)){
  company = founder_leads_clean$company_id[i]
  if(prev_company == company){
    l = l + 1
  }else{
    prev_company = company
    company_index = i
    l = 0
  }
  if(l < nleads){
    founders_bg[company_index,l*3+1] = founder_leads_clean$degree_type[i]
    founders_bg[company_index,l*3+2] = founder_leads_clean$subject[i]
    founders_bg[company_index,l*3+3] = founder_leads_clean$institution[i]
    if(l == 0) founders_bg[company_index,7] = length(founder_leads_clean$company_id[founder_leads_clean$company_id == company & founder_leads_clean$title == "Founder"])
  }
}

#View(founders_bg)
founders_clean = cbind(founder_leads_clean, founders_bg)
founders_clean = founders_clean[founders_clean$leader1_degree_type != "0",c("company_id", "degree_type", "subject", "institution", "title")]
founders_clean = na.omit(founders_clean)
#View(founders_clean)
```


## Investment & Aquisition Data 

```{r}

invest_rounds = invest_rounds[,c("company_id", "funding_round_code", "raised_amount_usd", "pre_money_valuation_usd", "post_money_valuation_usd", "participants")]
#View(invest_rounds)
str(invest_rounds)

rounds_table = as.data.table(invest_rounds)
rounds_table = melt(data = rounds_table, id.vars=c("company_id", "funding_round_code"), measure.vars = c("raised_amount_usd", "pre_money_valuation_usd", "post_money_valuation_usd", "participants"), variable_name = "value")
invest_rounds = na.omit(as.data.frame(dcast(data = rounds_table, formula = company_id ~ funding_round_code + variable, fun.aggregate = sum)))
#View(rounds_table)
#str(invest_rounds)

```

```{r}
aquisitions = aquisitions[aquisitions$price_currency_code == "USD",c("company_id", "aquisition_price_amount")]
nrow(aquisitions)
```


# Final Dataset For Analysis

```{r}
company_data = merge(companies, aquisitions, by = "company_id")
nrow(company_data)
company_data = merge(company_data, founders_clean, by = "company_id")
nrow(company_data)
company_data = merge(company_data, invest_rounds, by = "company_id")
nrow(company_data)

str(company_data)
View(company_data)
```


