---
title: "Data Analysis - Project Proposal"
author: "Mohamed Diakite Pineda, Yu Liu, Diego Carreno, STAT 420, Summer 2020"
date: "7/19/2020"
output:
  html_document: 
    theme: readable
    toc: yes
  pdf_document: default
urlcolor: cyan
---

***

## Student names

- Mohamed Diakite Pineda, *md21@*
- Yu Liu, *yul10@*
- Diego Carreno, *diegoac3@*

## Project Title

Building a predictive model for startup invesments and valuation

## Description of the data file

Pending.

## Background information of the data set

The data set comes from Kaggle (see [source](https://www.kaggle.com/justinas/startup-investments?select=funding_rounds.csv)), and it contains diverse data from the startup environment in 2013. It includes, but is not limited to:

- Organizations
- Founders information
- Rounds and funding dates
- Acquisitions
- Valuation

## Purpose statement

We are interested in predicting startup success and valuation given a combination of factors, such as number of founders, rounds of funding, number of aquisitions, founders background, and location among others. We want to explore and see if there are specific predictor variables from which we can build a model that can explain startup success and/or valuation reasonably well.

## Loading the data into `R`

---
title: "Startup Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Data load 

The startup data comes from Kaggle: [https://www.kaggle.com/justinas/startup-investments]

This dataset contains information about the startup ecosystem: organizations, individuals, company news, funding rounds, acquisitions, and IPOs, extracted originally from the Crunchabse Data website. In this project, only 6 of the 11 tables will be used and  joined using unique IDs. 

Bellow is the loading of the data.

```{r message=FALSE}

library(readr)

#People
people = read_csv("./people.csv")
(peep_n = colnames(people))
peep_n[2] = "person_id"
colnames(people) = peep_n
attr(people, "spec") = NULL
str(people)

```

```{r message=FALSE}
#Company
companies = read_csv("./objects.csv")
#(comp_n = colnames(companies))
attr(companies, "spec") = NULL
attr(companies, "problems") = NULL
colnames(companies)[1] = "company_id"
str(companies)
```

```{r message=FALSE}

degrees = read_csv("./degrees.csv")
colnames(degrees)
colnames(degrees)[2] = "person_id"
colnames(degrees)
attr(degrees, "spec") = NULL
str(degrees)

```


```{r message=FALSE}
comp_peep = read_csv("./relationships.csv")
colnames(comp_peep)[3] = "person_id"
colnames(comp_peep)[4] = "company_id"
colnames(comp_peep)
attr(comp_peep, "spec") = NULL
str(comp_peep)
```


```{r message=FALSE}
#install.packages("reshape")
library(reshape)
library(data.table)

invest_rounds = read_csv("./funding_rounds.csv")
attr(invest_rounds, "spec") = NULL
colnames(invest_rounds)[3]="company_id"
str(invest_rounds)

```


```{r message=FALSE}
aquisitions = read_csv("./acquisitions.csv")
colnames(aquisitions)[4] = "company_id"
colnames(aquisitions)[6] = "aquisition_price_amount"

attr(aquisitions, "spec") = NULL
str(aquisitions)

#unique(aquisitions$price_currency_code)
#aggregate(aquisitions[aquisitions$price_currency_code == "USD",c("price_currency_code", "company_id")], by = list(aquisitions$price_currency_code), FUN = counts)#aquisitions = 

```


```{r}
length(aquisitions$aquisition_price_amount[aquisitions$aquisition_price_amount>0])
```

### Data Preparation

#### Company Founders


This joins the people, degrees and their relationship to the companies. Also,this section "normalizes" the data, particularly the discreet data, so it can be used for regression.

```{r}
full_peep = merge(people, degrees, by="person_id")
str(full_peep)
```

```{r }
#join people with comp_peeps
founders = merge(full_peep, comp_peep, by="person_id")
str(founders)
#View(founders)
```

```{r}
#Identify founders

#get rows with founder, ceos, chairmans only
leads = c(grep("founder", ignore.case = TRUE, x = founders$title), grep("ceo", ignore.case = TRUE, x = founders$title), grep("chairman", ignore.case = TRUE, x = founders$title))
founders[grep("founder", ignore.case = TRUE, x = founders$title), "title"] = "Founder"
founder_leads = founders[leads, ]
#nrow(founder_leads)

#Functions to coubt rows for analysis of data
count_disticts = function(x){
  length(unique(x))
}

counts = function(x){
  length(x)
}
```


"Normalize" the title information for the founders' degree type so that it can be used as a potential regression predictor.

```{r}
founder_leads_clean = founder_leads
founder_leads_clean[grep("MS ", ignore.case = TRUE, x = founder_leads_clean$degree_type), c("degree_type")] = "MS"
founder_leads_clean[grep("Master in Business|Master of Business|M\\.B\\.A\\.|MBA|Executive", ignore.case = TRUE, x = founder_leads_clean$degree_type), c("degree_type")] = "MBA"
founder_leads_clean[grep("Ph|doc|pos.*grad", ignore.case = TRUE, x = founder_leads_clean$degree_type), c("degree_type")] = "Ph.D."
founder_leads_clean[grep("Bachelor of Science|BSc|B\\.S|B\\.E|BE|BS|Ing|Eng|Bsc|SCB", ignore.case = TRUE, x = founder_leads_clean$degree_type), c("degree_type")] = "BS"
founder_leads_clean[grep("Bachelor|BA|B\\.A", ignore.case = TRUE, x = founder_leads_clean$degree_type), c("degree_type")] = "BA"
founder_leads_clean[grep("B[A-Z][A-Z]+|B\\.|BC", ignore.case = FALSE, x = founder_leads_clean$degree_type), c("degree_type")] = "Bachelors"
founder_leads_clean[grep("MS|M\\.S", ignore.case = TRUE, x = founder_leads_clean$degree_type), c("degree_type")] = "MS"
founder_leads_clean[grep("Master|Graduate|Fellow", ignore.case = TRUE, x = founder_leads_clean$degree_type), c("degree_type")] = "Masters"
founder_leads_clean[grep("M[A-Z]+|M.", ignore.case = FALSE, x = founder_leads_clean$degree_type), c("degree_type")] = "Masters"
founder_leads_clean[grep("No.*degre|high.*school|secondary educa|no.*complete|course|incomplete|none|did not|prepa|[0-9].*[0-9]+ Grade", ignore.case = TRUE, x = founder_leads_clean$degree_type), c("degree_type")] = "No Degree"
founder_leads_clean[grep("Law|JD|J\\.D\\.", ignore.case = FALSE, x = founder_leads_clean$degree_type), c("degree_type")] = "Law"
founder_leads_clean[grep("Business", ignore.case = TRUE, x = founder_leads_clean$degree_type), c("degree_type")] = "MBA"
founder_leads_clean[grep("Undergr|Diplom|degree|college|under grad|lic|dipl", ignore.case = TRUE, x = founder_leads_clean$degree_type), c("degree_type")] = "Bachelors"
founder_leads_clean[!(founder_leads_clean$degree_type %in% c("MS", "MBA", "Ph.D.", "BS", "BA", "Bachelors", "MS", "Masters", "No Degree", "Law", "MBA")), c("degree_type")] = "Degree"

#View(unique(founder_leads_clean$degree_type))
analysis = aggregate(founder_leads_clean[,c("degree_type", "person_id", "company_id")], by = list(founder_leads_clean$degree_type), FUN = counts)


#analysis[order(-analysis$person_id), ]
#analysis[order(analysis$person_id)]
founder_leads_clean = na.omit(founder_leads_clean[,c("degree_type", "subject", "institution", "company_id", "title")])
founder_leads_clean = founder_leads_clean[order(founder_leads_clean$company_id),]
#View(founder_leads_clean)

```

Reshape the people subdataset into a wider table that has founder information per company, each in a row of data. This transposes the founder information (education background) to a single row on a per-company basis.

```{r}

founders_bg = cbind(
  leader1_degree_type = rep(0, nrow(founder_leads_clean)), leader1_subject = rep(0, nrow(founder_leads_clean)), leader1_institution = rep(0, nrow(founder_leads_clean)), 
  leader2_degree_type = rep(0, nrow(founder_leads_clean)), leader2_subject = rep(0, nrow(founder_leads_clean)), leader2_institution = rep(0, nrow(founder_leads_clean)), 
  founders_ceos_chairmans = rep(0, nrow(founder_leads_clean)))

nleads = 2
prev_company = "c"
company_index = 0
l = 0
for(i in  1:nrow(founder_leads_clean)){
  company = founder_leads_clean$company_id[i]
  if(prev_company == company){
    l = l + 1
  }else{
    prev_company = company
    company_index = i
    l = 0
  }
  if(l < nleads){
    founders_bg[company_index,l*3+1] = founder_leads_clean$degree_type[i]
    founders_bg[company_index,l*3+2] = founder_leads_clean$subject[i]
    founders_bg[company_index,l*3+3] = founder_leads_clean$institution[i]
    if(l == 0) founders_bg[company_index,7] = length(founder_leads_clean$company_id[founder_leads_clean$company_id == company & founder_leads_clean$title == "Founder"])
  }
}

```

REmove additional rows and columns that are relevant for the regression

```{r}

#View(founders_bg)
founders_clean = cbind(founder_leads_clean, as.data.frame(founders_bg))
founders_clean = founders_clean[founders_clean$leader1_degree_type != "0",-5]
founders_clean = na.omit(founders_clean)
#View(founders_clean)

```


#### Investment & Aquisition Data 

Investment data is stored in "long" shape, with every company having multiple rounds of investment (Series A-F, debt, angel, seed, etc.), each stored in a single row. Below is the process of reshaping the inventment rounds to a wide shape, in which every row represents all the investment rounds for a single company.


```{r}

invest_rounds = invest_rounds[,c("company_id", "funding_round_code", "raised_amount_usd", "pre_money_valuation_usd", "post_money_valuation_usd", "participants")]
#View(invest_rounds)
str(invest_rounds)


rounds_table = as.data.table(invest_rounds)
rounds_table = melt(data = rounds_table, id.vars=c("company_id", "funding_round_code"), measure.vars = c("raised_amount_usd", "pre_money_valuation_usd", "post_money_valuation_usd", "participants"), variable_name = "value")
invest_rounds = na.omit(as.data.frame(dcast(data = rounds_table, formula = company_id ~ funding_round_code + variable, fun.aggregate = sum)))
#View(rounds_table)
#str(invest_rounds)

```

Filter out companies with aquisition amount not in USD (all other variable/predictors are in USD)
```{r}
aquisitions = aquisitions[aquisitions$price_currency_code == "USD",c("company_id", "aquisition_price_amount")]

```


### Final Dataset For Analysis

After extracting the individual data aspects of companies, these aspects are joined into a single dataset. Depending on the resulting row numbers, we will decide which dataset will be used for the linear regression.

```{r}
company_aq = company_data = merge(companies, aquisitions, by = "company_id")
nrow(company_aq)
company_inv = merge(company_data, invest_rounds, by = "company_id")
nrow(company_inv)
company_aq_lds = merge(company_data, founders_clean, by = "company_id")
nrow(company_aq_lds)
str(company_aq)
```

### Preminiary Analysis

```{r}
startups = company_aq[company_aq$aquisition_price_amount>0, ]
write_csv(x = startups, path = "./startups.csv")
nrow(startups)

#Linear Model
company_model = lm(log(aquisition_price_amount) ~ city+ log(funding_total_usd+1) + category_code + funding_rounds + milestones + relationships, data=startups)

#R^2 of model
summary(company_model)$r.squared

#F statistic and p-value of model
fstat = summary(company_model)$fstatistic
pf(fstat[1], fstat[2], fstat[3], lower.tail=FALSE)

#Plots of the model
par(mfrow = c(1,2))
plot(company_model, which = 1:2)



```


